NumPy---Array Computation for Python

### 2020-01-06 body.tex
Switch figures

Data analysis frequently involves operations $(+, -, \times, /, \log, \sin, \dots)$ on
the individual measurements.
In programming languages such as C or Java, this would involve looping over
every element to perform these operations.
Array computations generalize these element-wise operations on scalars to
vectors (1-d arrays), matrices (2-d arrays), and higher dimensional arrays (n-d
arrays).
For instance, if \code{x} is an array, then \code{2 * x} produces
an array whose values are twice those in \code{x}.
Similarly, if \code{x} and \code{y} have the same shape, then
\code{x * y} produces an array whose values are the products of
the corresponding entries of \code{x} and \code{y}.
When \code{x} and \code{y} have different shapes, the smaller array is \emph{broadcast}
across the larger array to a common shape when sensible.
These vectorized operations often produce one-liners that would be many tens
of lines of code in languages such as C.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{scripts/PRs-using-CURRENT_MAINTAINERS.pdf}
    \caption{Number of pull requests merged into the NumPy master branch for each
        quarter since 2012. The total number of PRs is indicated with the
        lower blue area showing the portion contributed by current or previous
        maintainers.}\label{fig:prs-over-time}
\end{figure}

NumPy is currently maintained by a group of 23 contributors with commit rights
to the NumPy code base. Out of these, 17 maintainers were active in
2019, 4 of whom were paid to work on the project full-time.
Additionally, there are a few long term developers who contributed and maintain
specific parts of NumPy, but are not officially maintainers.
At a release cycle of about every half year, the five recent releases in the years
2018 and 2019 have averaged about 450~pull requests~(PRs) each,\footnote{
    Note that before mid 2011, NumPy development did not happen on \url{github.com}.
    All data provided here is based on the development which happened through GitHub
    pull requests. In some cases contributions by maintainers may not be categorized as such.}
% Since 1.14.0 (based on changelog): 381 + 438 + 490 + 531 + 402; last is preliminary
with each release attracting more than a hundred new contributors.
Figure~\ref{fig:prs-over-time} shows the number of PRs merged into the NumPy
master branch.
Although the number of PRs being merged fluctuates,
the plot indicates an increased number of contributions over the past
years.

>>>>

Data analysis frequently involves operations $(+, -, \times, /, \log, \sin, \dots)$ on
the individual measurements.
In programming languages such as C or Java, this would involve looping over
every element to perform these operations.
Array computations generalize these element-wise operations on scalars to
vectors (1-d arrays), matrices (2-d arrays), and higher dimensional arrays (n-d
arrays).
For instance, if \code{x} is an array, then \code{2 * x} produces
an array whose values are twice those in \code{x}.
Similarly, if \code{x} and \code{y} have the same shape, then
\code{x * y} produces an array whose values are the products of
the corresponding entries of \code{x} and \code{y}.
When \code{x} and \code{y} have different shapes, the smaller array is \emph{broadcast}
across the larger array to a common shape when sensible.
For example, if \code{x} is a 1-d array with length 3, and \code{y} is a 2-d
array with shape $2 \times 1$, the output will be an array with shape $2 \times 3$
(see Figure~\ref{fig:broadcasting}).

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{static/broadcasting}
  \caption{
    Broadcasting a $1 \times 3$ array against a $2 \times 1$ array
    yields a $2 \times 3$ result.
  }
  \label{fig:broadcasting}
\end{figure}

These vectorized operations often produce one-liners that would be many tens
of lines of code in languages such as C.

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.9\linewidth]{scripts/PRs-using-CURRENT_MAINTAINERS.pdf}
%    \caption{Number of pull requests merged into the NumPy master branch for each
%        quarter since 2012. The total number of PRs is indicated with the
%        lower blue area showing the portion contributed by current or previous
%        maintainers.}\label{fig:prs-over-time}
%\end{figure}
%
%NumPy is currently maintained by a group of 23 contributors with commit rights
%to the NumPy code base. Out of these, 17 maintainers were active in
%2019, 4 of whom were paid to work on the project full-time.
%Additionally, there are a few long term developers who contributed and maintain
%specific parts of NumPy, but are not officially maintainers.
%At a release cycle of about every half year, the five recent releases in the years
%2018 and 2019 have averaged about 450~pull requests~(PRs) each,\footnote{
%    Note that before mid 2011, NumPy development did not happen on \url{github.com}.
%    All data provided here is based on the development which happened through GitHub
%    pull requests. In some cases contributions by maintainers may not be categorized as such.}
%% Since 1.14.0 (based on changelog): 381 + 438 + 490 + 531 + 402; last is preliminary
%with each release attracting more than a hundred new contributors.
%Figure~\ref{fig:prs-over-time} shows the number of PRs merged into the NumPy
%master branch.
%Although the number of PRs being merged fluctuates,
%the plot indicates an increased number of contributions over the past
%years.
